---
title: "Portfolio 4"
---

In this project, I will use the dataset diamonds."A dataset containing the prices and other attributes of almost 54,000 diamonds." The goal is to fit a linear regression model to examine how carat explains the variation in diamond prices. 

load the data
```{r}
library(ggplot2)
data(diamonds)
head(diamonds)
summary(diamonds)
?diamonds
```

Visualizing Price Distribution
```{r}
ggplot(diamonds, aes(x = price)) +
  geom_histogram(binwidth = 200, fill = "purple",color = "black", alpha = 0.5) +
  labs(title = "Distribution of Diamond Prices", x = "Price (USD)", y = "Count")
```
This graph shows a right-skewed distribution, with most diamond prices falling below $5000. 

Check the Correlation 
```{r}
cor(diamonds[, c("price", "carat", "depth", "table", "x", "y", "z")])

```

Scatter Plot: Price vs Carat
```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Diamond Price vs Carat Weight", x = "Carat", y = "Price")

```

Price, Carat, Linear Regression Model
```{r}
Price_Carat <- lm(price ~ carat, data=diamonds)
summary (Price_Carat)
summary(Price_Carat)$r.squared
```

Intercept -2256.36
Slope 7756.43 
Price  = 7756.43 carat -2256.36
R squared = 0.8493305
Slope 7756.43 . This means that for every 1 unit increase in carat weight, the price of the diamond increases by $7756.43 on average.
Intercept -2256.36. This represents the predicted price of a diamond when carat = 0. However, since diamonds with zero carat weight do not exist, this value is not meaningful in practice. 
R squared = 0.8493305. means that approximately 84.93% of the variance in diamond prices is explained by the carat variable alone. 

