---
title: "Portfolio 6"
---

In this portfolio, I will continue the project started at portfolio 4 and 5, using the dataset diamonds. I will fit a linear regression model to see how the clarity and color of a diamond affects its price and how much it explains the differences in prices. 
 
load the data
```{r}
library(ggplot2)
data(diamonds)
head(diamonds)
summary(diamonds)
?diamonds
```

Price and color
```{r}
ggplot(diamonds, aes(x = color, y = price)) +
  stat_summary(fun = mean, geom = "bar", fill = "skyblue", color = "black") +
  labs(title = "Average Diamond Price by color", x = "color", y = "Average Price")

```

Price and clarity 
```{r}
ggplot(diamonds, aes(x = clarity, y = price)) +
  stat_summary(fun = mean, geom = "bar", fill = "skyblue", color = "black") +
  labs(title = "Average Diamond Price by Clarity", x = "Clarity", y = "Average Price")

```

Price, color, Linear Regression Model
```{r}
# Convert 'color' to an unordered factor
diamonds$color <- factor(diamonds$color, ordered = FALSE)

Price_Color <- lm(price ~ color, data=diamonds)
summary (Price_Color)
summary(Price_Color)$r.squared
```

Intercept 3169.95
Price  = -93.20 colorE  + 554.93 colorF + 829.18 colorG + 1316.72 colorH +  1921.92 colorI +  2153.86 colorJ + 3169.95
R squared = 0.03127542
Adjusted R-squared:  0.03117 
R squared = 0.03127542 . means that approximately 3.13% of the variance in diamond prices is explained by the carat variable alone. 

Price, clarity , Linear Regression Model
```{r}
diamonds$clarity <- factor(diamonds$clarity, ordered = FALSE)
Price_Clarity <- lm(price ~ clarity, data=diamonds)
summary (Price_Clarity)
summary(Price_Clarity)$r.squared
```
Intercept 3924.1687
Price  = 1138.8599 claritySI2 + 71.8325 claritySI1 +  0.8207 clarityVS2 + -84.7133 clarityVS1 + -640.4316 clarityVVS2 -1401.0541 clarityVVS1 -1059.3296 clarityIF + 3924.1687

R squared = 0.0271503
Adjusted R-squared:  0.02702 
R squared = 0.0271503 . means that approximately 2.72% of the variance in diamond prices is explained by the carat variable alone. 

Price, color and clarity, Linear Regression Model
```{r}
Price_Clarity_Color <- lm(price ~ clarity+color, data=diamonds)
summary (Price_Clarity_Color)
summary(Price_Clarity_Color)$r.squared
```

R squared = 0.05937474
Adjusted R-squared:  0.05915, the explanatory power increased by adding the color factor. 
R squared = 0.05937474 . means that approximately 5.9% of the variance in diamond prices is explained by the carat variable alone. 


